---
title: "Modis Data Download"
author: "Justin Williams"
date: '2022-06-05'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  root.dir = "/Users/justinwilliams/projects/geotmseries")
```

```{r load-packages, warning=FALSE, message=FALSE}
if(!require(knitr)) install.packages("knitr", 
repos = "http://cran.us.r-project.org")
if(!require(tidyverse)) install.packages("tidyverse", 
repos = "http://cran.us.r-project.org")
if(!require(sf)) install.packages("sf", 
repos = "http://cran.us.r-project.org")
if(!require(terra)) install.packages("terra", 
repos = "http://cran.us.r-project.org")
if(!require(raster)) install.packages("raster", 
repos = "http://cran.us.r-project.org")
if(!require(mapview)) install.packages("mapview", 
repos = "http://cran.us.r-project.org")
if(!require(MODIStsp)) install.packages("MODIStsp", 
repos = "http://cran.us.r-project.org")
if(!require(dotenv)) install.packages("dotenv", 
repos = "http://cran.us.r-project.org")
if(!require(raster)) install.packages("raster", 
repos = "http://cran.us.r-project.org")
if(!require(rasterVis)) install.packages("rasterVis",
repos = "http://cran.us.r-project.org")
if(!require(DEPONS2R)) install.packages("DEPONS2R",
repos = "http://cran.us.r-project.org")
if(!require(ggOceanMaps)) install.packages("ggOceanMaps",
repos = "http://cran.us.r-project.org")
```

## US County shapefile

Here we will obtain US County Shapefile from [Census.gov]("https://www2.census.gov/geo/tiger/GENZ2018/shp/cb_2018_us_county_5m.zip") and utilize it to visualize previously agreed upon extent for MODIS data download.
This will help us visualize the area in question, and verify accuracy of information downloaded.

```{r county-shapefile}
# create tempfiles
temp <- tempfile()
temp_2 <- tempfile()

# download zipfile save to temp
URL <- "https://www2.census.gov/geo/tiger/GENZ2018/shp/cb_2018_us_county_5m.zip"
download.file(URL, temp)

# unzip
unzip(zipfile = temp, exdir = temp_2)

#read in shapefile
county_df <- st_read(temp_2)
```

Let's clip section to agreed upon extent bounding box (xmin, xmax, ymin, ymax).

```{r clip-bbox}
# clip to bbox
clipped_shape <- clip_shapefile(county_df,
               limits = c(-118.828125,-116.015625,35.01915128,37.82067492),
               proj.limits = convert_crs(4326),
               simplify = F,
               tol = 60,
               return.boundary = F)

# preview clip
mapview(clipped_shape)

# convert to sf
clipped_shape_sf <- st_as_sf(clipped_shape)
```

## Modis exploration

Get product names that are surface temperature related.

```{r get-product-code}
# surface temperature
surface_temp <- MODIStsp_get_prodnames()[13:20]

surface_temp
```

Let's loop through these and see what product layers are available for each one. 

```{r loop-through-products}
# extrapolate product name
surface_temp_prod_name <- surface_temp %>% 
  str_extract("(?<=\\().*(?=\\))")

# loop over each product name to see what layers are available
map(.x = surface_temp_prod_name,
    .f = MODIStsp_get_prodlayers)
```

## Modis data download

Set dropbox for storage off laptop

```{r dropbox_dir, eval=FALSE}
dropbox <- "C:\\Users\\jtws\\Dropbox\\data_projects\\GeoTimeSeries\\data_daily_7_4_22"
```

Download daily 1km `Surf_Temp_Daily_1Km (M*D11A1)`

```{r download-2021-monthly-005dg, eval=FALSE}
MODIStsp(gui             = FALSE,
         out_folder      = dropbox,
         out_folder_mod  = dropbox,
         selprod         = 'Surf_Temp_Daily_1Km (M*D11A1)',
         bandsel         = "LST_Day_1km", # daily surface temp
         sensor = "Terra",
         # your username for NASA http server
         user            = Sys.getenv("EARTH_DATA_USER") ,
         # your password for NASA http server
         password        = Sys.getenv("EARTH_DATA_PW"),
         start_date      = '2000.01.01', 
         end_date        = '2020.12.31', 
         verbose         = TRUE,
         spatmeth = "bbox",
         bbox = c(-118.828125,35.01915128,-116.015625,37.82067492),
         out_format      = 'GTiff',
         compress        = 'LZW',
         out_projsel     = 'User Defined',
         output_proj     = 4326,
         delete_hdf      = FALSE,
         parallel        = TRUE,
         reprocess = FALSE
         )

```

## Exploratory data analysis (EDA)

Now that data is downloaded, will conduct brief EDA to view exactly what was acquired.

Get files from 2015 - 2020.

```{r files-by-year}
# set path
# dropbox_path <- r"(C:\Users\jtws\Dropbox\data_projects\GeoTimeSeries\data_daily_7_4_22\Surf_Temp_Daily_1Km_v6\LST_Day_1km)"
dropbox_path <- "/Users/justinwilliams/Dropbox/data_projects/GeoTimeSeries/data_daily_7_4_22/Surf_Temp_Daily_1Km_v6/LST_Day_1km"

# function to list files from specific year
extract_year <- function(pattern, path = dropbox_path) {
  list.files(path = dropbox_path,
             pattern = pattern,
             full.names = T)
}

# create list
daily_sur_temp <- list()
for (i in 2015:2020) {
  daily_sur_temp <- append(daily_sur_temp,paste0("daily_sur_temp_",i))
}

# pattern for extract function
years <- as.character(c(2015:2020))

# map to list
daily_sur_temp_2015_2020 <- 
  map(.x = years,
      .f = extract_year)

# change list names to years
names(daily_sur_temp_2015_2020) <- daily_sur_temp
```

Convert to Spatraster, then brick, then stack, divided by year.

```{r convert-to-Spatraster-brick-stack}
# function to convert tiff files to stack
stack_brick_rast <- function(raster_files) {
  stack(brick(rast(raster_files)))
}

# map function
daily_sur_temp_2015_2020_stack <- 
  map(.x = daily_sur_temp_2015_2020,
    .f = stack_brick_rast)
```

Function to aggregate monthly.

```{r spot-check-missing-data}
monthly_agg_temp <- function(temp_stack, month_name, 
                              agg_function = "mean") {
  
  # format names and create indices
  indices <- format(
    as.Date(names(temp_stack), 
            format="MOD11A1_LST_Day_1km_%Y_%j"), 
            format = "%m")
  indices <- as.numeric(indices) # coerce to num
  
  # agg mean monthly
  month_agg <- stackApply(
        x = temp_stack,
        indices = indices,
        fun = agg_function,
        na.rm = T)
  
  # change names to month
  names(month_agg) <- month.name
  
  # subset raster for named month
  raster::subset(
        month_agg, 
        grep(month_name,names(month_agg),
        value = T))
}
```

Apply month agg function.

```{r map-month-agg-function}
mean_july_2015 <- monthly_agg_temp(daily_sur_temp_2015_stack,"July")

plot(mean_july_2015)

mean_tas_july_2015_2020 <- map(.x = daily_sur_temp_2015_2020_stack,
                           .f = monthly_agg_temp,
                           month_name = "July")

mean_july_2015 <- monthly_agg_temp(daily_sur_temp_2015_2020_stack$daily_sur_temp_2015, "July")



monthly_agg_temp(daily_sur_temp_2015_2020_stack$daily_sur_temp_2015,
                "July")
```


Scale Kelvin and convert to Fahrenheit.

```{r convert-temp}
convert_fah <- function(raster_stack) {
  # convert 0 to NA
  raster_stack[raster_stack == 0] <- NA
  # convert to Fahrenheit
  1.8 * (raster_stack * 0.02 - 273) + 32
}

# apply to each raster file
daily_sur_temp_2015_f <- convert_fah(daily_sur_temp_2015_stack)
daily_sur_temp_2016_f <- convert_fah(daily_sur_temp_2016_stack)
daily_sur_temp_2017_f <- convert_fah(daily_sur_temp_2017_stack)
daily_sur_temp_2018_f <- convert_fah(daily_sur_temp_2018_stack)
daily_sur_temp_2019_f <- convert_fah(daily_sur_temp_2019_stack)
daily_sur_temp_2020_f <- convert_fah(daily_sur_temp_2020_stack)
```


Convert to Brick. Multiple Raster objects can be combined to produce a stack or brick. These data objects are frequently used for space-time data because they can store multiple rasters from different dates. 

```{r tempbrick}
# create brick
tempbrick_2015_2020_f <- 
  brick(daily_sur_temp_2015_rast_f, daily_sur_temp_2016_rast_f,      daily_sur_temp_2017_rast_f, daily_sur_temp_2018_rast_f, daily_sur_temp_2019_rast_f, daily_sur_temp_2020_rast_f)

# name years
names(tempbrick_2015_2020_f) <- 
  c("Temp.2015","Temp.2016", "Temp.2017", 
    "Temp.2018", "Temp.2019", "Temp.2020")
```

Get cellstats of each layer

```{r cellstats}
cellStats(tempbrick_2015_2020_f, stat = "mean")
```

Function to convert raster to df.

```{r raster-to-df}
rasterdf <- function(x, aggregate = 1) {
  resampleFactor <- aggregate        
  inputRaster <- x    
  inCols <- ncol(inputRaster)
  inRows <- nrow(inputRaster)
  # Compute numbers of columns and rows in the new raster for mapping
  resampledRaster <- raster(ncol=(inCols / resampleFactor), 
                            nrow=(inRows / resampleFactor))
  # Match to the extent of the original raster
  extent(resampledRaster) <- extent(inputRaster)
  # Resample data on the new raster
  y <- resample(inputRaster,resampledRaster,method='ngb')
  # Extract cell coordinates into a data frame
  coords <- xyFromCell(y, seq_len(ncell(y)))
  # Extract layer names
  dat <- stack(as.data.frame(getValues(y)))
  # Add names - 'value' for data, 'variable' to indicate different raster layers
  # in a stack
  names(dat) <- c('value', 'variable')
  dat <- cbind(coords, dat)
  dat
}
```

Apply `rasterdf()` function to brick.

```{r raster-to-df-brick}
tempbrick_df <- rasterdf(tempbrick_2015_2020_f)
summary(tempbrick_df)
```

Plot df with **ggplot**.

```{r plot-df}
# plot temp years
ggplot() +
  geom_raster(data = tempbrick_df, 
              aes(x = x, y = y, fill = value)) +
  geom_sf(data=clipped_shape_sf,
          fill=NA,color="grey25", size=0.50) +
  scale_fill_gradient(name = "Degrees F", 
                      low = "yellow", high = "red") +
  facet_wrap(~ variable, ncol = 3) + 
  labs(title = "Mean Temperature (F) 2015 - 2020") +
  theme_void() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5))
```

Why does 2018 have so little records?

```{r plot-2018}
daily_sur_temp_2018_rast_f_df <- rasterdf(daily_sur_temp_2018_rast_f)
daily_sur_temp_2018_rast_f_df %>% 
  ggplot() + geom_raster(aes(x, y, fill = value))

plot(daily_sur_temp_2018_rast)

sum(!is.na(daily_sur_temp_2018_rast_f_df$value))
```

Looks like there is a bunch of missing data. This will have to be imputed at a
latter stage. Let's look at a comparsion  off

Histogram of brick

```{r hist-brick}
hist(tempbrick_2015_2020_f)
```

Density of brick

```{r density-brick}
density(tempbrick_2015_2020_f)
```

Density plot

```{r density-plot}
densityplot(tempbrick_2015_2020_f)
```

Compare years with correlation coefficient

```{r correlation}
pairs(tempbrick_2015_2020_f)
```

Let's compare temperature anomalies across these 6 years.

```{r temp-anomaly}
# get mean
meantemp <- mean(tempbrick_2015_2020_f, na.rm = T)
# get tempanom
tempanom <- tempbrick_2015_2020_f - meantemp
# set names
names(tempanom) <- names(tempbrick_2015_2020_f)
# turn into df
tempanom_df <- rasterdf(tempanom)

# plot
ggplot() +
  geom_raster(data = tempanom_df, aes(x = x, y = y, fill = value)) +
  geom_sf(data=clipped_shape_sf, fill=NA,color="grey25", size=0.50) +
  scale_fill_gradient2(name = "Degrees F", 
                       low = "blue", 
                       mid = "lightyellow", high = "red") +
  facet_wrap(~ variable, ncol = 3) + 
  labs(title = "Mean Temperature Orange County 2015 - 2020 (F)") +
  theme_void() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5))

```

## Resources

[MODIStsp package documentation](https://docs.ropensci.org/MODIStsp/)
[Geospatial Data Science with R](https://bookdown.org/mcwimberly/gdswr-book/raster-data-continuous-variables.html)

